The language of generalization: Probability, vagueness, and interaction

### 2018-02-18 paper/coverletter.tex
minor tweaks to cl

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

@@

We enclose the manuscript \emph{The Language of Generalization} to be considered for publication in \emph{Psychological Review}. 
This paper explores the linguistic means by which generalizations are communicated.  
We present a Bayesian model of endorsing generalizations and show that it predicts context-sensitive human judgments with high quantitative accuracy across several experiments in diverse domains.
The theory suggests that the language of generalization can be understood in terms of a simple semantic theory that operates over beliefs, represented as probability.

Understanding how the world works requires learning generalizations. 
The kinds of observations that would facilitate such learning may be rare, however (e.g., figuring out that lightning strikes tall objects) or costly (e.g., learning that certain plants are poisonous).
We do not need to need taste every plant to figure out which ones are poisonous, however.
Language lets us learn generalizations from each other.

The language of generalization is most commonly studied in predications of categories (e.g.,~``Birds fly.''), otherwise known as \emph{generic language}.
But generalization exists beyond classically defined categories (e.g., natural and artifact kinds): Events are described in generalization (e.g.,~``Mary swims after work.''; \emph{habitual language}) as well as causal relationships (e.g.,~``Staring at the sun makes you go blind''; \emph{causal language}).
Generalization are ubiquitous in everyday conversation (``Movie theaters are cold.''), the speech of young children (``Slides aren't for grown ups!''), child-directed speech (``Dogs are friendly.''), political and scientific discourse (``Politicians make promises.'', ``Psychology experiments don't replicate.''), stereotypes (``Boys are good at math.''), motivation (``Big girls eat broccoli!''), and many other facets of experience.

@@

We have previously received peer-review of a related manuscript submitted to top-ranking journals including \emph{Psychological Review}.
Those reviews raised a umber of points regarding the presentation and generality of our work at the time, and highlighted the need for comparison to other models. 
The current manuscript differs from previous submissions in a number of important ways.
First, we have completely re-written the paper for clarity and focus, especially simplifying and streamlining the modeling sections.
Second, we have extended from generalizations about categories (or \emph{generic language}) to investigate three case studies: generalizations about categories, events, and causes.
We are able to explain endorsements across these three domains with minimal domain-specific assumptions. 
Furthermore, each case study naturally lends methodological advantages to provide further tests of our hypothesis/model (e.g., with causal generalizations, we are able to manipulate---rather than merely measure---prior knowledge).
%The empirical and much of the philosophical literatures have focused on generalizations about categories, or generic language.
%Our previous manuscript also focused on categories; in this paper, however, we investigate three case studies: generalizations about categories, events, and causes. 
%Immediately we see the generality of our formal model: We are able to explain context-sensitive endorsements across these three domains without adopting domain-specific assumptions.
%Furthermore, each case study naturally lends methodological advantages to provide further tests of our model.
%With event generalizations, we are able to interrogate the underlying semantic scale, and find that speakers' subjective beliefs about what is likely to be the case in the future, and not the objective truth about what has been the case in the past, is what drives endorsements. 
%With causal generalizations, we are able to \emph{manipulate} (rather than merely measure) prior knowledge, and show it is in fact causally related to endorsement. 
Finally, we develop three alternative models: two linear models based on empirically-elicited statistical knowledge and one rich, cognitive model which lacks the \emph{underspecification of meaning} at the core of our hypothesis. 
These alternative models fail to account for key aspects of the context-sensitivity that is observed in the data and that our full model predicts.
%These alternative models and the empirical paradigms used provide strong tests of the theoretical contribution of our paper. 
%This novel approach for generalizations in language opens up a number of interesting questions about the nature of conceptual structure and its linguistic traces, which we discuss at the end of the paper.
Overall we feel that this paper represents a much improved, generalized and streamlined investigation of generalization in language.
(We would be happy to provide detailed responses to previous reviewers; presently, we feel this would be of limited use given the major changes in content and presentation.)

@@

This work will be of interest to psychologists, as well as linguists, philosophers, and computer scientists interested in language and interaction.
The puzzles of generalizations in language are immediately compelling to a general audience, the more so because of their connections to cognitive development, propagation of stereotypes, and cultural transmission of knowledge.
Our solution represents the first formal model of generalization in language to be validated by quantitative agreement with human behavioral data.
It is not, however, uncontroversial: we champion a statistical view that has been often dismissed in the literature.
Indeed, in presentations around the world we have found this work to excite broad interest and engagement.

>>>>

\textwidth 6.55in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

@@

We enclose the manuscript \emph{The Language of Generalization} to be considered for publication in \emph{Psychological Review}. 
This paper explores the linguistic means by which generalizations are communicated.  
We present a Bayesian model of endorsing generalizations and show that it predicts context-sensitive human judgments with high quantitative accuracy across diverse experimental contexts.
The theory suggests that the language of generalization can be understood in terms of a simple semantic theory that operates over beliefs, represented as probability.

Understanding how the world works requires learning generalizations. 
The kinds of observations that would facilitate such learning may be rare, however (e.g., figuring out that lightning strikes tall objects) or costly (e.g., learning that certain plants are poisonous).
We do not need to need taste every plant to figure out which ones are poisonous, however.
Language lets us learn complex and abstract generalizations from each other.

The language of generalization is most commonly studied in predications of categories (e.g.,~``Birds fly.''), otherwise known as \emph{generic language}.
But generalization exists beyond classically-defined natural and artifact kinds: Events are described in generalization (e.g.,~``Mary swims after work.''; \emph{habitual language}) as well as causal relationships (e.g.,~``Staring at the sun makes you go blind''; \emph{causal language}).
Generalizations are ubiquitous in everyday conversation (``Movie theaters are cold.''), the speech of young children (``Slides aren't for grown ups!''), child-directed speech (``Dogs are friendly.''), political and scientific discourse (``Politicians make promises.'', ``Psychology experiments don't replicate.''), motivation (``Big girls eat broccoli!''), stereotypes (``Boys are good at math.''), and many other facets of experience.

@@

We have previously received peer-review of a related manuscript submitted to top-ranking journals including \emph{Psychological Review}.
Those reviews raised a number of points regarding the presentation and generality of our work at the time, and highlighted the need for comparison to other models. 
The current manuscript differs from previous submissions in a number of important ways.
First, we have completely re-written the paper for clarity and focus, especially simplifying and streamlining the modeling sections.
Second, we have extended from generalizations about categories (or \emph{generic language}) to investigate three case studies: generalizations about categories, events, and causes.
We are able to explain endorsements across these three domains with minimal domain-specific assumptions. 
Furthermore, each case study naturally lends methodological advantages to provide further tests of our hypothesis/model (e.g., with causal generalizations, we are able to manipulate---rather than merely measure---prior knowledge).
%The empirical and much of the philosophical literatures have focused on generalizations about categories, or generic language.
%Our previous manuscript also focused on categories; in this paper, however, we investigate three case studies: generalizations about categories, events, and causes. 
%Immediately we see the generality of our formal model: We are able to explain context-sensitive endorsements across these three domains without adopting domain-specific assumptions.
%Furthermore, each case study naturally lends methodological advantages to provide further tests of our model.
%With event generalizations, we are able to interrogate the underlying semantic scale, and find that speakers' subjective beliefs about what is likely to be the case in the future, and not the objective truth about what has been the case in the past, is what drives endorsements. 
%With causal generalizations, we are able to \emph{manipulate} (rather than merely measure) prior knowledge, and show it is in fact causally related to endorsement. 
Finally, we develop three alternative models: two linear models based on empirically-elicited statistical knowledge and one rich, cognitive model which lacks the \emph{underspecification of meaning} at the core of our hypothesis. 
These alternative models fail to account for key aspects of the context-sensitivity that is observed in the data and that our full model predicts.
%These alternative models and the empirical paradigms used provide strong tests of the theoretical contribution of our paper. 
%This novel approach for generalizations in language opens up a number of interesting questions about the nature of conceptual structure and its linguistic traces, which we discuss at the end of the paper.
Overall we feel that this paper represents a much improved, generalized and streamlined investigation of generalization in language, and a major breakthrough in computational modeling of language understanding.
(We would be happy to provide detailed responses to previous reviewers; presently, we feel this would be of limited use given the major changes in content and presentation.)

@@

This work will be of interest to psychologists, as well as linguists, philosophers, and computer scientists interested in language and interaction.
The puzzles of generalizations in language are immediately compelling to a general audience, the more so because of their connections to cognitive development, propagation of stereotypes, and cultural transmission of knowledge.
Our solution represents the first formal model of generalization in language to be validated by quantitative agreement with human behavioral data.
It is not, however, uncontroversial: We champion a statistical view that has been often dismissed in the literature.
Indeed, in presentations around the world we have found this work to excite broad interest and engagement.
